### 2012-11-23 paper/paper.tex
Paper v1

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
   \usepackage{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{C:\Users\Administrator\Pictures}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
   \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Low-level routing for quality optimization of scalable video streaming on the future Internet}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Piet Smet, Sander Vrijders, Youri Flement, Bruno Chevalier}
\IEEEauthorblockA{University of Ghent - Computer Science, Ghent, Belgium}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
In current IP networks, routes are usually based on the hop count from sender to receiver without keeping in mind the throughput and loss of packets. IP is a best-effort protocol and doesn't guarantee zero packet loss. This is unsatisfactory if we want to deliver qualitative, scalable video streams. This paper presents a framework and algorithms used to provide Quality of Service (\textit{QoS}) support for scalable video streaming in an OpenFlow network. We aim to solve two problems: first we route the base layer of SVC encoded videos as a lossless-QoS flow while the enhancement layers are routed as a lossy-QoS or best effort flow. Second, we provide failure recovery so the OpenFlow network can converge back to an optimal state when links go down. We show that using different costs to calculate paths for QoS flows can improve the overall throughput of video streams when network convergation and delay occurs.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
In a standarized IP network, each router or switch contains a flow table containing the next hop for each destination, often based on the shortest path calculation. The network devices do not know what data a packet contains and they do not care either, each flow follows the same rules. The Internet Protocol (IP) headers contain a priority field but this is often neglected by most routers. Media applications, such as video streaming, require no packet loss and minimal network delay due to their high Quality of Service (QoS) dependancy. When these QoS flows are sent through the same path as other best effort traffic they are subject to network congestion and delays which renders the outcome unsatisfactory.
Scalable Video Coding (SVC) encodes the stream in a base layer along with several enhancement layers. In order to obtain correct output it is important that the base layer has no packet loss and no network delay, something that can't be guaranteed if there are no seperate forwarding rules set up for this QoS flow. To solve this issue we need the routing algorithms to recognize QoS flows and seperate them from normal best effort traffic.\\In this project we are looking to use OpenFlow in order to provide QoS support for scalable video streaming. To achieve this goal we can route the base layer of SVC encoded video as a lossless-QoS flow while the enhancement layers are routed as a lossy-QoS or best effort flow, as described in. Even though this approach deviates from the shortest path algorithms, we can still show that alternative paths can increase the video’s throughput in case of network congestion. We have also provided means of failure recovery so the OpenFlow network c
 an conve
 rge back to an optimal state when links go down.\\OpenFlow [1] is an Internet architecture developed by Stanford University that enables researchers to run experimental protocols in existing networks with heterogeneous switches. This is achieved by decoupling the control and forwarding layers of routing, where each flow is associated with an action. With OpenFlow, vendors do not need to expose the internal working of their products as OpenFlow exploits a common set of functions present in most switches to achieve its goal. Switches forward packages of new flows to a centralized Controller, after which the Controller can set up the routing tables of switches in the network. Listener modules on the Controller can use filters to seperate our QoS flows from other traffic.\\In Section II of this paper we compare related research to our own. In section III we take a closer look at the components of our framework and the principals it is based on. In section IV we present the diffe
 rent alg
 orithms used to determine the behavior of each flow in the network. Section V shows the test scenario we used to evaluate our framework while Section VI presents the found results.

\begin{figure*}[t]
\centering
\includegraphics[width=\linewidth]{SVC_OpenFlow_flow}
\caption{Simulation Results}
\label{fig_sim}
\end{figure*}

\section{Related work}
In our research we identified a need for seperate forwarding actions per flow in order apply load-balancing. The same principal has been discussed in research papers from Princeton University [2]. In their research the aim was to apply load-balancing during service discovery of web services when all requests had to be processed by a central component. A problem occurs when each request is forwarded to a centralized component responsible for service discovery, which leads to network congestion. By using OpenFlow, the researchers at Princeton University managed to set seperate forwarding actions for client requests which match certain criteria. These actions allow forwarding entities to redirect client requests to an appropriate service instance without interference from the centralized component. Even though the use case may be different, the principal is the same as the one applied in our research. When both best effort and QoS flows follow the same paths, chances for congest
 ion and 
 packet loss to occur increase significantly. OpenFlow allows us to set different paths for both flows, which improves the throughput of our QoS flows and also decrease the network load on links used by best effort traffic.\\The idea to provide a minimal quality for SVC encoded videos originates from research described in [4] and [5]. By routing the base layer as a lossless-QoS flow we can guarantee a minimal quality to the user. Ideally, the enhancement layers will also be delivered without packet loss or delay but when congestion occurs we demote the enhancement layers to best effort traffic and prioritise the base layer. Research discussed in [4] also attempts to route the enhancement layers as a lossy-QoS flow if the network allows it.\\The OpenFlow Controller becomes a possible bottleneck for wide-area networks if too many QoS flows need to be processed for a first time. HyperFlow [3] suggests using areas across the network, each containing atleast one Controller. The co
 ntroller
 s of different areas can then exchange information about their areas to obtain a global overview. In our research we work with relatively small networks so we assume that one Controller is sufficient.

\section{Architecture}
Each OpenFlow network contains atleast one Controller which is responsible for managing all the switches, this is no different in our research. In this project we used Floodlight Controller to implement our Java modules, which contain all non-default code. A default OpenFlow switch forwards packages to the Controller if the package does not match any forwarding entry. We study SVC encoded video streams as QoS flow, where we consider the base layer as a lossless-QoS flow and the enhancement layer as best effort traffic, when congestion occurs. These principals are shown in figure 1. A filter on the Controller seperates QoS flows from other traffic, after which the Controller can decide if the shortest path is still sufficient to carry the base layer. When the package is part of a QoS flow and congestion occurs on the shortest path, we calculate an alternative path using metrics other than just the hop count.\\This was one of the first tasks we performed during our research: to
  find me
 trics, besides the hop count, which can be used to calculate the cost of a path to transfer zero packet loss QoS traffic. These metrics are part of a cost function which the routing algorithm will attempt to minimize when calculating optimal paths. We made similar assumptions about QoS-routing as those made in [4] and [5] before defining a cost function:
\begin{itemize}
	\item When lossless-QoS flows are forwarded on the same link as best effort traffic, there will be packet loss in the best effort flow to prioritise the QoS flow.
	\item Metrics for QoS flows must be different than those used for best effort traffic. The shortest path and QoS path must be different to get decent results. If both follow the same path, the best effort traffic will be completely sacrificed for the sake of the lossless-QoS base layer.
	\item Calculating a path for a new QoS flow does not change the best effort routing tables. The packet loss of best effort traffic caused by QoS flows must be considered during path selection.
	\item The length of alternative paths for QoS flows should still be kept as short as possible to optimize transfer time.
	\item Best effort traffic can be used as a reference to retreive metrics from a switch or link, such as packet loss or to estimate the amount of traffic on that route.
\end{itemize}
These assumptions indicate that packet loss and path length are part of our cost function: the cost function must find a balance between maintaining low packet loss of the best effort traffic and the delay of QoS flows, caused by using longer paths.\\Furthermore, the Controller must be able to obtain these metrics at any time from any switch in the network before calculating new flow tables. This is why we implemented a monitor module on the Controller to gather all network statistics. Besides polling for statistics, a switch must also forward events to the Controller when they occur, such as an adjacent link going down or a flow experiencing packet loss. These events provide additional monitoring data for the routing algorithms on the Controller.\\Decisions made by the Controller are sent to the switches as updated forwarding tables. When the shortest path can still ensure zero packet loss, the Dijkstra shortest path algorithm is executed, which is the default implementation
  in Floo
 dlight.

\section{QoS Routing}
Must ensure zero packet loss base layer, optimise other traffic ... cost function?
Go from theoretical cost function to exact function in code?

\section{Evaluation}
Test scenario (client/sever - filter based on port) + metrics for result + present results\\
An evaluation metric has to represent a value which is relevant to what we want to improve, which is why a measurement must:
\begin{itemize}
	\item represent the quality of the received video
	\item represent the performance of the network
	\item represent the impact on best effort traffic
\end{itemize}
To measure quality of the video we have chosen for the PSNR value of the received video compared to the raw footage. The higher the PSNR value, the better the frames were preserved during transfer. To measure the performance of the network we used the bitrate of all traffic. We also show the packet loss of best effort traffic caused by prioritising the QoS flow...

\section{Conclusion}
The conclusion goes here.

%\subsection{Subsection Heading Here}
%Subsection text here\cite{IEEEexample:confwithpaper}.

%\subsubsection{Subsubsection Heading Here}
%Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

% conference papers do not normally have an appendix

% use section* for acknowledgement
%\section*{Acknowledgment}
%The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,IEEEexample}

% that's all folks
\end{document}

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/videofilter/VideoFilter.java

@@ -6,8 +6,11 @@ import java.util.Arrays;
 import java.util.Collection;
 import java.util.Comparator;
 import java.util.EnumSet;
+import java.util.HashMap;
+import java.util.LinkedList;
 import java.util.List;
 import java.util.Map;
+import java.util.Set;
 
 import net.floodlightcontroller.core.FloodlightContext;
 import net.floodlightcontroller.core.IFloodlightProviderService;
@@ -25,10 +28,15 @@ import net.floodlightcontroller.devicemanager.SwitchPort;
 import net.floodlightcontroller.packet.Ethernet;
 import net.floodlightcontroller.packet.IPv4;
 import net.floodlightcontroller.packet.TCP;
+import net.floodlightcontroller.routing.BroadcastTree;
 import net.floodlightcontroller.routing.IRoutingDecision;
+import net.floodlightcontroller.routing.Link;
 import net.floodlightcontroller.routing.Route;
+import net.floodlightcontroller.routing.RouteId;
+import net.floodlightcontroller.topology.Cluster;
 import net.floodlightcontroller.topology.ITopologyService;
 import net.floodlightcontroller.topology.NodePortTuple;
+import net.floodlightcontroller.topology.TopologyInstance;
 import net.floodlightcontroller.util.OFMessageDamper;
 
 import org.openflow.protocol.OFFlowMod;
@@ -109,7 +117,6 @@ public class VideoFilter implements IOFMessageListener, IFloodlightModule {
 	@Override
 	public void startUp(FloodlightModuleContext context) {
 		floodlightProvider.addOFMessageListener(OFType.PACKET_IN, this);
-		floodlightProvider.addOFMessageListener(OFType.STATS_REPLY, this);
 
 	}
 
@@ -136,6 +143,7 @@ public class VideoFilter implements IOFMessageListener, IFloodlightModule {
 
 					// TODO: Call doForwardFlow with the calculated weights
 					// change doForwardFlow args...
+					doForwardFlow(sw, pi, cntx);
 				}
 
 				// Get the TCP port of the packet
@@ -248,7 +256,12 @@ public class VideoFilter implements IOFMessageListener, IFloodlightModule {
 
 						// TODO: INSERT DIJKSTRA HERE FOR OUR VIDEO FLOWS
 						// AND CALCULATE FLOWS
-						Route route;
+						// Call getroute also with weights!!
+
+						Route route = getRoute(srcDap.getSwitchDPID(),
+								(short) srcDap.getPort(),
[…]

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/forwarding/Constants.java
Changes

@@ -1,0 +1,9 @@
+package net.floodlightcontroller.forwarding;
+
+public class Constants {
+
+	public static final int BASE = 8080;
+	public static final int ENHANCE1 = 8081;
+	public static final int ENHANCE2 = 8082;
+	public static final int ENHANCE3 = 8083;
+}

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/forwarding/Forwarding.java

@@ -244,11 +244,25 @@ public class Forwarding extends ForwardingBase implements IFloodlightModule {
                     if (!srcDap.equals(dstDap) && 
                         (srcCluster != null) && 
                         (dstCluster != null)) {
-                        Route route = 
-                                routingEngine.getRoute(srcDap.getSwitchDPID(),
-                                                       (short)srcDap.getPort(),
-                                                       dstDap.getSwitchDPID(),
-                                                       (short)dstDap.getPort());
+                    	
+                    	
+                    	Route route = routingEngine.getRoute(srcDap.getSwitchDPID(),
+                                (short)srcDap.getPort(),
+                                dstDap.getSwitchDPID(),
+                                (short)dstDap.getPort());
+                    	if(match.getTransportSource() == Constants.BASE){
+                    		//Re-request route from routing engine but with weights this time
+                    		
+                    		
+                    		
+                    	} else if (match.getTransportSource() == Constants.ENHANCE1){
+                    		
+                    	} else if (match.getTransportSource() == Constants.ENHANCE2){
+                    		
+                    	} else if (match.getTransportSource() == Constants.ENHANCE3){
+                    		
+                    	}
+                    	
                         if (route != null) {
                             if (log.isTraceEnabled()) {
                                 log.trace("pushRoute match={} route={} " + 

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/videofilter/VideoFilter.java

@@ -143,23 +143,10 @@ public class VideoFilter implements IOFMessageListener, IFloodlightModule {
 
 					// TODO: Call doForwardFlow with the calculated weights
 					// change doForwardFlow args...
-					doForwardFlow(sw, pi, cntx);
+					//doForwardFlow(sw, pi, cntx);
 				}
 
-				// Get the TCP port of the packet
-				TCP tcppacket = (TCP) ippacket.getPayload();
-				int port = tcppacket.getSourcePort();
-
-				if (port == Constants.BASE) {
-
-				} else if (port == Constants.ENHANCE1) {
-
-				} else if (port == Constants.ENHANCE2) {
-
-				} else if (port == Constants.ENHANCE3) {
-
-				}
-
+				
 			}
 		}
 		return Command.CONTINUE;
@@ -592,42 +579,42 @@ public class VideoFilter implements IOFMessageListener, IFloodlightModule {
     
     
     protected Route buildroute(RouteId id, long srcId, long dstId) {
-    	
-    	Map<Long, BroadcastTree> destinationRootedTrees = new HashMap<Long, BroadcastTree>();
-    	Map<Link, Integer> linkCost = new HashMap<Link, Integer>();
-    	Set<Cluster> clusters = topology
-    	
-    	
-    	
-    	for(Cluster c: clusters) {
-            for (Long node : c.links.keySet()) {
-                BroadcastTree tree = dijkstra(c, node, linkCost, true);
-                destinationRootedTrees.put(node, tree);
-            }
-        }
-    	
+//    	
+//    	Map<Long, BroadcastTree> destinationRootedTrees = new HashMap<Long, BroadcastTree>();
+//    	Map<Link, Integer> linkCost = new HashMap<Link, Integer>();
+//    	Set<Cluster> clusters = topology
+//    	
+//    	
+//    	
+//    	for(Cluster c: clusters) {
+//            for (Long node : c.links.keySet()) {
+//                BroadcastTree tree = dijkstra(c, node, linkCost, true);
+//                destinationRootedTrees.put(node, tree);
+//            }
+//        }
+//    	
     	
         NodePortTuple npt;
 
         LinkedList<NodePortTuple> switchPorts =
                 new LinkedList<NodePortTuple>();
 
-        if (destinationRootedTrees.get(dstId) == null) return null;
+//        if (destinationRootedTrees.get(dstId) == null) return null;
+//
+//        Map<Long, Link> nexthoplinks =
+//                destinationRootedTrees.get(dstId).getLinks();
[…]

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/forwarding/Forwarding.java
Updaye

@@ -250,17 +250,12 @@ public class Forwarding extends ForwardingBase implements IFloodlightModule {
                                 (short)srcDap.getPort(),
                                 dstDap.getSwitchDPID(),
                                 (short)dstDap.getPort());
-                    	if(match.getTransportSource() == Constants.BASE){
+                    	if(match.getTransportSource() == Constants.BASE || match.getTransportSource() == Constants.ENHANCE1 || match.getTransportSource() == Constants.ENHANCE2 || match.getTransportSource() == Constants.ENHANCE3){
                     		//Re-request route from routing engine but with weights this time
-                    		
-                    		
-                    		
-                    	} else if (match.getTransportSource() == Constants.ENHANCE1){
-                    		
-                    	} else if (match.getTransportSource() == Constants.ENHANCE2){
-                    		
-                    	} else if (match.getTransportSource() == Constants.ENHANCE3){
-                    		
+                    		route = routingEngine.getRoute(srcDap.getSwitchDPID(),
+                                    (short)srcDap.getPort(),
+                                    dstDap.getSwitchDPID(),
+                                    (short)dstDap.getPort(), (int)match.getTransportSource());
                     	}
                     	
                         if (route != null) {

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/routing/IRoutingService.java

@@ -36,7 +36,10 @@ public interface IRoutingService extends IFloodlightService {
     public Route getRoute(long srcId, short srcPort, 
                              long dstId, short dstPort, 
                              boolean tunnelEnabled);
-
+    public Route getRoute(long srcId, short srcPort, 
+                             long dstId, short dstPort, int tcpport);
+    
+    
     /** Check if a route exists between src and dst, including tunnel links
      *  in the path.
      */

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/topology/TopologyInstance.java

@@ -15,6 +15,8 @@ import org.slf4j.Logger;
 import org.slf4j.LoggerFactory;
 
 import net.floodlightcontroller.util.ClusterDFS;
+import net.floodlightcontroller.core.IOFSwitch;
+import net.floodlightcontroller.core.NetworkStatistics;
 import net.floodlightcontroller.core.annotations.LogMessageCategory;
 import net.floodlightcontroller.core.annotations.LogMessageDoc;
 import net.floodlightcontroller.routing.BroadcastTree;
@@ -584,12 +586,18 @@ public class TopologyInstance {
         }
     }
 
-    protected Route buildroute(RouteId id, long srcId, long dstId) {
+    protected Route buildroute(RouteId id, long srcId, long dstId, int tcpport) {
         NodePortTuple npt;
 
         LinkedList<NodePortTuple> switchPorts =
                 new LinkedList<NodePortTuple>();
-
+        
+        //TODO: Construct a map with link weights
+        Map<IOFSwitch, Long> load = NetworkStatistics.networkLoad();
+        
+        //TODO: Re-call calculateShortestPathTreeInCluster with link weights
+        
+        
         if (destinationRootedTrees == null) return null;
         if (destinationRootedTrees.get(dstId) == null) return null;
 
@@ -624,6 +632,9 @@ public class TopologyInstance {
         if (log.isTraceEnabled()) {
             log.trace("buildroute: {}", result);
         }
+        
+        calculateShortestPathTreeInClusters();
+        
         return result;
     }
 
@@ -651,7 +662,7 @@ public class TopologyInstance {
     }
 
     protected Route getRoute(long srcId, short srcPort,
-                             long dstId, short dstPort) {
+                             long dstId, short dstPort, int tcpport) {
 
 
         // Return null the route source and desitnation are the
@@ -661,7 +672,7 @@ public class TopologyInstance {
 
         List<NodePortTuple> nptList;
         NodePortTuple npt;
-        Route r = getRoute(srcId, dstId);
+        Route r = getRoute(srcId, dstId, tcpport);
         if (r == null && srcId != dstId) return null;
 
         if (r != null) {
@@ -679,13 +690,13 @@ public class TopologyInstance {
         return r;
     }
 
-    protected Route getRoute(long srcId, long dstId) {
+    protected Route getRoute(long srcId, long dstId, int tcpport) {
         RouteId id = new RouteId(srcId, dstId);
         Route result = null;
         if (pathcache.containsKey(id)) {
[…]

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/topology/TopologyManager.java

@@ -515,7 +515,7 @@ public class TopologyManager implements
     @Override
     public Route getRoute(long src, long dst, boolean tunnelEnabled) {
         TopologyInstance ti = getCurrentInstance(tunnelEnabled);
-        return ti.getRoute(src, dst);
+        return ti.getRoute(src, dst, 0);
     }
 
     @Override
@@ -526,10 +526,14 @@ public class TopologyManager implements
     @Override
     public Route getRoute(long src, short srcPort, long dst, short dstPort, 
                           boolean tunnelEnabled) {
-        TopologyInstance ti = getCurrentInstance(tunnelEnabled);
-        return ti.getRoute(src, srcPort, dst, dstPort);
+        return getRoute(src, srcPort, dst, dstPort, 0);
     }
 
+    public Route getRoute(long src, short srcPort, long dst, short dstPort, int tcpport) {
+    	 TopologyInstance ti = getCurrentInstance(true);
+         return ti.getRoute(src, srcPort, dst, dstPort, tcpport);
+    }
+    
     @Override
     public boolean routeExists(long src, long dst) {
         return routeExists(src, dst, true);

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/topology/TopologyInstance.java
Update

@@ -1,6 +1,7 @@
 package net.floodlightcontroller.topology;
 
 import java.util.ArrayList;
+import java.util.Collection;
 import java.util.HashMap;
 import java.util.HashSet;
 import java.util.Iterator;
@@ -10,7 +11,7 @@ import java.util.Map;
 import java.util.PriorityQueue;
 import java.util.Set;
 
-
+import org.openflow.protocol.OFPhysicalPort;
 import org.slf4j.Logger;
 import org.slf4j.LoggerFactory;
 
@@ -26,808 +27,879 @@ import net.floodlightcontroller.routing.RouteId;
 import net.floodlightcontroller.util.LRUHashMap;
 
 /**
- * A representation of a network topology.  Used internally by 
+ * A representation of a network topology. Used internally by
  * {@link TopologyManager}
  */
 @LogMessageCategory("Network Topology")
 public class TopologyInstance {
 
-    public static final short LT_SH_LINK = 1;
-    public static final short LT_BD_LINK = 2;
-    public static final short LT_TUNNEL  = 3; 
-
-    public static final int MAX_LINK_WEIGHT = 10000;
-    public static final int MAX_PATH_WEIGHT = Integer.MAX_VALUE - MAX_LINK_WEIGHT - 1;
-    public static final int PATH_CACHE_SIZE = 1000;
-
-    protected static Logger log = LoggerFactory.getLogger(TopologyInstance.class);
-
-    protected Map<Long, Set<Short>> switchPorts; // Set of ports for each switch
-    /** Set of switch ports that are marked as blocked.  A set of blocked
-     * switch ports may be provided at the time of instantiation. In addition,
-     * we may add additional ports to this set.
-     */
-    protected Set<NodePortTuple> blockedPorts;  
-    protected Map<NodePortTuple, Set<Link>> switchPortLinks; // Set of links organized by node port tuple
-    /** Set of links that are blocked. */
-    protected Set<Link> blockedLinks;
-
-    protected Set<Long> switches;
-    protected Set<NodePortTuple> broadcastDomainPorts;
-    protected Set<NodePortTuple> tunnelPorts;
-
-    protected Set<Cluster> clusters;  // set of openflow domains
-    protected Map<Long, Cluster> switchClusterMap; // switch to OF domain map
-
-    // States for routing
-    protected Map<Long, BroadcastTree> destinationRootedTrees;
[…]

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/topology/TopologyInstance.java
Update

@@ -651,7 +651,19 @@ public class TopologyInstance {
 			Map<Link, Integer> linkCost = new HashMap<Link, Integer>();
 
 			if (tcpport == 8080) {
-				
+				for (NodePortTuple nt : tunnelPorts) {
+					if (switchPortLinks.get(npt) == null)
+						continue;
+					for (Link link : switchPortLinks.get(npt)) {
+						if (link == null)
+							continue;
+						
+						//Here you have the link, now add a weight
+						
+						
+						linkCost.put(link, weight);
+					}
+				}
 			} else if (tcpport == 8081) {
 
 			} else if (tcpport == 8082) {

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/topology/TopologyInstance.java
Update

@@ -652,14 +652,21 @@ public class TopologyInstance {
 
 			if (tcpport == 8080) {
 				for (NodePortTuple nt : tunnelPorts) {
-					if (switchPortLinks.get(npt) == null)
+					if (switchPortLinks.get(nt) == null)
 						continue;
-					for (Link link : switchPortLinks.get(npt)) {
+					for (Link link : switchPortLinks.get(nt)) {
 						if (link == null)
 							continue;
 						
 						//Here you have the link, now add a weight
+						long src = link.getSrc();
+						long dst = link.getDst();
 						
+						long srcLoad = load.get(src);
+						long dstLoad = load.get(dst);
+						
+						//SAFE?
+						int weight = (int) (srcLoad + dstLoad);
 						
 						linkCost.put(link, weight);
 					}

### 2012-11-23 floodlight/src/main/java/net/floodlightcontroller/topology/TopologyInstance.java
Update

@@ -1,7 +1,6 @@
 package net.floodlightcontroller.topology;
 
 import java.util.ArrayList;
-import java.util.Collection;
 import java.util.HashMap;
 import java.util.HashSet;
 import java.util.Iterator;
@@ -11,7 +10,6 @@ import java.util.Map;
 import java.util.PriorityQueue;
 import java.util.Set;
 
-import org.openflow.protocol.OFPhysicalPort;
 import org.slf4j.Logger;
 import org.slf4j.LoggerFactory;
 

### 2012-12-03 floodlight/src/main/java/net/floodlightcontroller/topology/TopologyInstance.java
Update

@@ -664,7 +664,7 @@ public class TopologyInstance {
 						long dstLoad = load.get(dst);
 
 						// SAFE?
-						int weight = (int) (srcLoad + dstLoad);
+						int weight = (int) (srcLoad + dstLoad) + 1;
 
 						linkCost.put(link, weight);
 					}
@@ -685,7 +685,7 @@ public class TopologyInstance {
 						long dstLoad = load.get(dst);
 
 						// SAFE?
-						int weight = (int) (srcLoad + dstLoad);
+						int weight = (int) (srcLoad + dstLoad) + 1;
 
 						linkCost.put(link, weight);
 					}
@@ -706,7 +706,7 @@ public class TopologyInstance {
 						long dstLoad = load.get(dst);
 
 						// SAFE?
-						int weight = (int) (srcLoad + dstLoad);
+						int weight = (int) (srcLoad + dstLoad) +1;
 
 						linkCost.put(link, weight);
 					}
@@ -727,7 +727,7 @@ public class TopologyInstance {
 						long dstLoad = load.get(dst);
 
 						// SAFE?
-						int weight = (int) (srcLoad + dstLoad);
+						int weight = (int) (srcLoad + dstLoad) +1;
 
 						linkCost.put(link, weight);
 					}

### 2012-12-03 floodlight/src/main/java/net/floodlightcontroller/core/NetworkStatistics.java
kleine aanpassing aan methode

@@ -103,7 +103,7 @@ public class NetworkStatistics extends TimerTask {
 		    	}				    	
 		    	
 		    	currentLoad.put(sw, newload / (period / 1000));
-		    	//System.out.println("Load for switch: " + transmittedBytes.get(sw) + " bytes/second");
+		    	//System.out.println("Load for switch: " + currentLoad.get(sw) + " bytes/second");
 		    }
 		}
 	}
@@ -115,6 +115,7 @@ public class NetworkStatistics extends TimerTask {
 	public static synchronized Map<IOFSwitch, Long> networkLoad(){
 		//return the whole map with loads
 		if(!currentLoad.isEmpty()){
+			//System.out.println("Loads in map:" + currentLoad.size());
 			return currentLoad;
 		} else {
 			return null;
@@ -129,7 +130,8 @@ public class NetworkStatistics extends TimerTask {
 	public static synchronized Long switchLoad(IOFSwitch sw){	
 		//lookup the load in the map
 		if(currentLoad != null && currentLoad.containsKey(sw)){
-			return loadHistory.get(sw);
+			//System.out.println("Load for: " + sw + " is: " + loadHistory.get(sw));
+			return currentLoad.get(sw);
 		} else {
 			return null;
 		}

### 2012-12-04 floodlight/src/main/java/net/floodlightcontroller/core/NetworkStatistics.java
opvragen networkload adhv switch id

@@ -31,10 +31,10 @@ public class NetworkStatistics extends TimerTask {
 	protected int period;
 	
 	//a map with the previous loads of switches in the network
-	protected static Map<IOFSwitch, Long> loadHistory;
+	protected static Map<Long, Long> loadHistory;
 	
 	//a map that keeps record of the current load
-	protected static Map<IOFSwitch, Long> currentLoad;
+	protected static Map<Long, Long> currentLoad;
 	
 	/**
 	 * Creates a new statistics task
@@ -44,8 +44,8 @@ public class NetworkStatistics extends TimerTask {
 	public NetworkStatistics(IFloodlightProviderService floodlightProvider, int period){
 		this.floodlightProvider = floodlightProvider;
 		this.period = period;
-		loadHistory = new HashMap<IOFSwitch, Long>();
-		currentLoad = new HashMap<IOFSwitch, Long>();
+		loadHistory = new HashMap<Long, Long>();
+		currentLoad = new HashMap<Long, Long>();
 	}
 	
 	/**
@@ -92,17 +92,17 @@ public class NetworkStatistics extends TimerTask {
 		    	 
 		    	//determine the new load
 		    	long newload = 0;		    	
-		    	if(currentLoad.containsKey(sw)){
+		    	if(currentLoad.containsKey(sw.getId())){
 		    		//we've seen the switch before
-		    		newload = reply.getTransmitBytes() - loadHistory.get(sw);
-			    	loadHistory.put(sw, reply.getTransmitBytes());  		
+		    		newload = reply.getTransmitBytes() - loadHistory.get(sw.getId());
+			    	loadHistory.put(sw.getId(), reply.getTransmitBytes());  		
 		    	} else {
 		    		//this is the first time we see the switch
 		    		newload = reply.getTransmitBytes();		
-		    		loadHistory.put(sw,newload);
+		    		loadHistory.put(sw.getId(),newload);
 		    	}				    	
 		    	
-		    	currentLoad.put(sw, newload / (period / 1000));
+		    	currentLoad.put(sw.getId(), newload / (period / 1000));
 		    	//System.out.println("Load for switch: " + currentLoad.get(sw) + " bytes/second");
 		    }
 		}
@@ -112,7 +112,7 @@ public class NetworkStatistics extends TimerTask {
 	 * Determines the load of all the switches in the network
 	 * @return a map with loads of all switches or null if the network has no switches
 	 */
-	public static synchronized Map<IOFSwitch, Long> networkLoad(){
[…]

### 2012-12-04 floodlight/src/main/java/net/floodlightcontroller/topology/TopologyInstance.java

@@ -645,7 +645,8 @@ public class TopologyInstance {
 
 		if (tcpport != 0) {
 			// TODO: Construct a map with link weights
-			Map<IOFSwitch, Long> load = NetworkStatistics.networkLoad();
+			//indexen via switch ID
+			Map<Long, Long> load = NetworkStatistics.networkLoad();
 			Map<Link, Integer> linkCost = new HashMap<Link, Integer>();
 
 			if (tcpport == 8080) {

### 2012-12-04 floodlight/src/main/java/net/floodlightcontroller/topology/TopologyInstance.java
mogelijks eerste werkende versie algoritme

@@ -18,6 +18,7 @@ import net.floodlightcontroller.core.IOFSwitch;
 import net.floodlightcontroller.core.NetworkStatistics;
 import net.floodlightcontroller.core.annotations.LogMessageCategory;
 import net.floodlightcontroller.core.annotations.LogMessageDoc;
+import net.floodlightcontroller.devicemanager.SwitchPort;
 import net.floodlightcontroller.routing.BroadcastTree;
 import net.floodlightcontroller.routing.Link;
 import net.floodlightcontroller.routing.Route;
@@ -643,7 +644,7 @@ public class TopologyInstance {
 
 		LinkedList<NodePortTuple> switchPorts = new LinkedList<NodePortTuple>();
 
-		if (tcpport != 0) {
+	//	if (tcpport != 0) {
 			// TODO: Construct a map with link weights
 			//indexen via switch ID
 			Map<Long, Long> load = NetworkStatistics.networkLoad();
@@ -712,11 +713,12 @@ public class TopologyInstance {
 						linkCost.put(link, weight);
 					}
 				}
-			} else if (tcpport == 8083) {
-				for (NodePortTuple nt : tunnelPorts) {
-					if (switchPortLinks.get(nt) == null)
-						continue;
-					for (Link link : switchPortLinks.get(nt)) {
+			//} else if (tcpport == 8083) {
+			} else { 
+				
+				
+				for (Set<Link> linkset: switchPortLinks.values()) {
+					for (Link link : linkset) {
 						if (link == null)
 							continue;
 
@@ -729,7 +731,8 @@ public class TopologyInstance {
 
 						// SAFE?
 						int weight = (int) (srcLoad + dstLoad) +1;
-
+						log.debug("{}  {}",src,dst);
+						log.debug("{} \n",weight);
 						linkCost.put(link, weight);
 					}
 				}
@@ -737,7 +740,7 @@ public class TopologyInstance {
 
 			// Re-call calculateShortestPathTreeInCluster with link weights
 			calculateShortestPathTreeInClusters(linkCost);
-		}
+	//	}
 		if (destinationRootedTrees == null)
 			return null;
 		if (destinationRootedTrees.get(dstId) == null)
@@ -774,7 +777,21 @@ public class TopologyInstance {
 		if (log.isTraceEnabled()) {
 			log.trace("buildroute: {}", result);
 		}
+		
+		for (NodePortTuple nt : result.getPath()) {
+			if (switchPortLinks.get(nt) == null)
[…]

### 2012-12-06 paper/paper.tex
Paper

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Low-level routing for quality optimization of scalable video streaming on the future Internet}
+\title{Layer based routing to obtain quality optimization of scalable video streaming in the Future Internet}
 
@@
 \author{\IEEEauthorblockN{Piet Smet, Sander Vrijders, Youri Flement, Bruno Chevalier}
-\IEEEauthorblockA{University of Ghent - Computer Science, Ghent, Belgium}
+\IEEEauthorblockA{Ghent University - iMinds, Department of Information Technology,\\ 
+Gaston Crommenlaan 8 bus 201, 9050 Gent, Belgium}
 }
@@
 %\boldmath
-In current IP networks, routes are usually based on the hop count from sender to receiver without keeping in mind the throughput and loss of packets.
-IP is a best-effort protocol and doesn't guarantee zero packet loss.
-This is unsatisfactory if we want to deliver qualitative, scalable video streams.
-This paper presents a framework and algorithms used to provide Quality of Service (\textit{QoS}) support for scalable video streaming in an OpenFlow network.
-We aim to solve two problems: first we route the base layer of SVC encoded videos as a lossless-QoS flow while the enhancement layers are routed as a lossy-QoS or best effort flow.
-Second, we provide failure recovery so the OpenFlow network can converge back to an optimal state when links go down.
-We show that using different costs to calculate paths for QoS flows can improve the overall throughput of video streams when network convergation and delay occurs.
+In TCP/IP networks, there is no inherent support for Quality of Service.
+Routing is done on the minimum hop count from sender to receiver, and all traffic is best-effort traffic. 
+This is unsatisfactory for applications such as video streaming. 
+This paper presents an OpenFlow controller application which provides 
+Quality of Service (\textit{QoS}) support for Scalable Video Streaming (\textit{SVC}) in an OpenFlow network. 
+First we route the base layer of SVC encoded videos as a 
+high priority QoS flow while the enhancement layers are routed at a lower QoS or as a best-effort flow. 
+Second, we provide failure recovery so the OpenFlow network can converge back to an optimal 
+state when links fail. We show that using different costs to calculate paths for different flows 
+can improve the overall throughput of video streams when network congestion and delay occur.
 \end{abstract}
@@
 \section{Introduction}
-In a standarized IP network, each router or switch contains a flow table containing the next hop for each destination, often based on the shortest path calculation.
-The network devices do not know what data a packet contains and they do not care either, each flow follows the same rules.
-The Internet Protocol (IP) headers contain a priority field but this is often neglected by most routers.
-Media applications, such as video streaming, require no packet loss and minimal network delay due to their high Quality of Service (QoS) dependancy.
-When these QoS flows are sent through the same path as other best effort traffic they are subject to network congestion and delays which renders the outcome unsatisfactory.
+In a TCP/IP-based network, each router or switch contains a flow table containing the 
+next hop for each destination, based on the shortest path calculation. 
+Media applications, such as video streaming, 
+require no packet loss and minimal network delay due to their high Quality of Service (QoS) dependency. 
+When these QoS flows are sent over the same path as other best effort traffic they 
+are subject to network congestion and delay which renders an unsatisfactory outcome. 
+Several efforts have been made to support QoS in a TCP/IP-based network, such as DiffServ and IntServ, but these 
+solutions only work well in managed networks.
 Scalable Video Coding (SVC) encodes the stream in a base layer along with several enhancement layers. 
-In order to obtain correct output it is important that the base layer has no packet loss and no network delay, something that can't be guaranteed if there are no seperate forwarding rules set up for this QoS flow.
-To solve this issue we need the routing algorithms to recognize QoS flows and seperate them from normal best effort traffic.\\In this project we are looking to use OpenFlow in order to provide QoS support for scalable video streaming.
-To achieve this goal we can route the base layer of SVC encoded video as a lossless-QoS flow while the enhancement layers are routed as a lossy-QoS or best effort flow, as described in.
-Even though this approach deviates from the shortest path algorithms, we can still show that alternative paths can increase the video’s throughput in case of network congestion.
-We have also provided means of failure recovery so the OpenFlow network c
- an conve
- rge back to an optimal state when links go down.\\OpenFlow [1] is an Internet architecture developed by Stanford University that enables researchers to run experimental protocols in existing networks with heterogeneous switches.
-This is achieved by decoupling the control and forwarding layers of routing, where each flow is associated with an action.
-With OpenFlow, vendors do not need to expose the internal working of their products as OpenFlow exploits a common set of functions present in most switches to achieve its goal.
-Switches forward packages of new flows to a centralized Controller, after which the Controller can set up the routing tables of switches in the network.
-Listener modules on the Controller can use filters to seperate our QoS flows from other traffic.\\In Section II of this paper we compare related research to our own.
-In section III we take a closer look at the components of our framework and the principals it is based on.
-In section IV we present the diffe
- rent alg
- orithms used to determine the behavior of each flow in the network. Section V shows the test scenario we used to evaluate our framework while Section VI presents the found results.
+In order to obtain correct output it is important that the base layer has no 
+packet loss and minimal delay, which is not possible without support for QoS. 
+To solve this issue we need the routing algorithms to detect QoS flows and route these flows differently than 
+normal best effort traffic.
+To achieve this goal in video streaming, we route the base layer of SVC encoded video at a high QoS 
+while the enhancement layers are routed either at a lower QoS or as a best effort flow. 
+We also provide failure recovery so the network 
+can converge back to an optimal state when links go down.\\
 
-\begin{figure*}[t]
+We use OpenFlow in order to provide QoS support for scalable video streaming. 
+OpenFlow is a protocol for switches, allowing to control the packet forwarding process.
+This way, networks are not only configurable, but also programmable.
+Each OpenFlow Switch has a flow table, which is used for packet lookup and 
+forwarding, and a secure channel to an external controller (see Figure \ref{fig:openflow1}). 
+The flow table consists of flow entries.
+Each flow table entry matches a traffic pattern, by matching the header fields.
+The header fields that can be matched in OpenFlow 1.0 can be seen in Table \ref{tab:matchedflows}.
+The controller makes the network programmable by managing the switch over 
+the secure channel using the OpenFlow protocol.
+The controller installs new entries in the flow table to manage the network traffic.
+Entries have a hard and a soft timeout. A soft timeout means the entry gets removed if it has not been used for a certain time.
+A hard timeout means the flow will be removed after the timeout, even if it is still in use.
+There are many implementations of controllers available for OpenFlow switches.
+The implementation of the controller we will use is Floodlight, because it is open source and widely used.
+
+\begin{table}[htb]
 \centering
-\includegraphics[width=\linewidth]{SVC_OpenFlow_flow}
+\begin{tabular}{| c | *{11}{c|}}
+\hline 
+In Port & Ethernet Source Address & Ethernet Destination Address \\ \hline
+Ethernet Type & VLAN ID & VLAN Priority \\ \hline
+IP Source & IP Destination & IP Protocol \\ \hline
+IP ToS Bits & TCP/UDP Source Port & TCP/UDP Destination Port\\ \hline 
+\end{tabular}
+\caption{Fields that can be matched in OpenFlow 1.0}
+\label{tab:matchedflows}
+\end{table}
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.6]{fig/openflow1}
+\caption{How OpenFlow works}
+\label{fig:openflow1}
+\end{figure}
+
+The rest of this paper is organized as follows.
+In Section II of this paper we compare related work to our own. 
+In section III we take a closer look at the components of our framework 
+and the principals it is based on. In section IV we present the different algorithms 
+used to determine the behavior of each flow in the network. Section V shows the test 
+scenario we used to evaluate our framework while Section VI presents results.
+In Section VII we come to a conclusion.
+
+\section{Related work}
+In our research we identified a need for separate forwarding actions per flow.
+The same principal has been discussed in research papers from Princeton University. 
+In their research the aim was to apply load-balancing during service discovery of web services 
+when all requests had to be processed by a central component. 
+A problem occurs when each request is forwarded to a centralized component responsible 
+for service discovery, which leads to network congestion. By using OpenFlow, the researchers at 
+Princeton University managed to set separate forwarding actions for client requests which match 
+certain criteria. These actions allow forwarding entities to redirect client requests to an 
+appropriate service instance without interference from the centralized component. Even though 
+the use case may be different, the principal is the same as the one applied in our research. 
+When both best effort and QoS flows follow the same paths, chances for congestion and packet loss 
+to occur increase significantly. OpenFlow allows us to set different paths for both flows, which 
+improves the throughput of our QoS flows and also decrease the network load on links used by best 
+effort traffic.\\
+
+The idea to provide a minimal quality for SVC encoded videos originates from 
+research described in [4] and [5]. By routing the base layer as a lossless-QoS flow we can 
+guarantee a minimal quality to the user. Ideally, the enhancement layers will also be delivered 
+without packet loss or delay but when congestion occurs we demote the enhancement layers to best 
+effort traffic and prioritize the base layer. Research discussed in [4] also attempts to route 
+the enhancement layers as a lossy-QoS flow. Our research goes further and also enables fast failure 
+recovery when a link fails.\\
+
+\section{Architecture}
+Each OpenFlow network contains at least one controller which is responsible for managing all the switches. 
+A default OpenFlow switch forwards packets 
+to the controller if the packet does not match any forwarding entry. 
+We route SVC encoded video streams as a QoS flow, where we consider the base layer as a 
+high priority QoS flow and the enhancement layer as either a lower QoS flow or as best effort traffic. 
+These principals are shown in Figure \ref{fig_sim}. A filter on the controller separates QoS flows from 
+other traffic, after which the controller can choose a path for the different layers. 
+When the controller calculates a path for a layer, the new forwarding rules are sent to the switches.\\
+
+\begin{figure*}[htb]
+\centering
+\includegraphics[width=\linewidth]{fig/SVC_OpenFlow_flow}
 \caption{Simulation Results}
@@
 
-\section{Related work}
-In our research we identified a need for seperate forwarding actions per flow in order apply load-balancing.
-The same principal has been discussed in research papers from Princeton University [2].
-In their research the aim was to apply load-balancing during service discovery of web services when all requests had to be processed by a central component.
-A problem occurs when each request is forwarded to a centralized component responsible for service discovery, which leads to network congestion.
-By using OpenFlow, the researchers at Princeton University managed to set seperate forwarding actions for client requests which match certain criteria.
-These actions allow forwarding entities to redirect client requests to an appropriate service instance without interference from the centralized component.
-Even though the use case may be different, the principal is the same as the one applied in our research.
-When both best effort and QoS flows follow the same paths, chances for congest
- ion and 
- packet loss to occur increase significantly.
-OpenFlow allows us to set different paths for both flows, which improves the throughput of our QoS flows and also decrease the network load on links used by best effort traffic.\\The idea to provide a minimal quality for SVC encoded videos originates from research described in [4] and [5].
-By routing the base layer as a lossless-QoS flow we can guarantee a minimal quality to the user.
-Ideally, the enhancement layers will also be delivered without packet loss or delay but when congestion occurs we demote the enhancement layers to best effort traffic and prioritise the base layer.
-Research discussed in [4] also attempts to route the enhancement layers as a lossy-QoS flow if the network allows it.\\The OpenFlow Controller becomes a possible bottleneck for wide-area networks if too many QoS flows need to be processed for a first time.
-HyperFlow [3] suggests using areas across the network, each containing atleast one Controller.
-The co
- ntroller
- s of different areas can then exchange information about their areas to obtain a global overview.
-In our research we work with relatively small networks so we assume that one Controller is sufficient.
-
-\section{Architecture}
-Each OpenFlow network contains atleast one Controller which is responsible for managing all the switches, this is no different in our research.
-In this project we used Floodlight Controller to implement our Java modules, which contain all non-default code.
-A default OpenFlow switch forwards packages to the Controller if the package does not match any forwarding entry.
-We study SVC encoded video streams as QoS flow, where we consider the base layer as a lossless-QoS flow and the enhancement layer as best effort traffic, when congestion occurs.
-These principals are shown in figure 1.
-A filter on the Controller seperates QoS flows from other traffic, after which the Controller can decide if the shortest path is still sufficient to carry the base layer.
-When the package is part of a QoS flow and congestion occurs on the shortest path, we calculate an alternative path using metrics other than just the hop count.\\This was one of the first tasks we performed during our research: to
-  find me
- trics, besides the hop count, which can be used to calculate the cost of a path to transfer zero packet loss QoS traffic.
-These metrics are part of a cost function which the routing algorithm will attempt to minimize when calculating optimal paths.
-We made similar assumptions about QoS-routing as those made in [4] and [5] before defining a cost function:
+We defined a cost function which can be used to calculate the cost of a path.
+The cost function is based on the following observations. (Insert ref here)
 \begin{itemize}
-	\item When lossless-QoS flows are forwarded on the same link as best effort traffic, there will be packet loss in the best effort flow to prioritise the QoS flow.
-	\item Metrics for QoS flows must be different than those used for best effort traffic.
-The shortest path and QoS path must be different to get decent results.
-If both follow the same path, the best effort traffic will be completely sacrificed for the sake of the lossless-QoS base layer.
+	\item When high priority QoS flows are forwarded on the same link as best effort traffic, there will be packet loss in the best effort flow to prioritize the QoS flow.
+	\item Metrics for QoS flows must be different than those used for best effort traffic. The shortest path and QoS path must be different to obtain good results. 
+	If both follow the same path, the best effort traffic will experience more packet loss.
 	\item Calculating a path for a new QoS flow does not change the best effort routing tables. The packet loss of best effort traffic caused by QoS flows must be considered during path selection.
 	\item The length of alternative paths for QoS flows should still be kept as short as possible to optimize transfer time.
-	\item Best effort traffic can be used as a reference to retreive metrics from a switch or link, such as packet loss or to estimate the amount of traffic on that route.
+	\item Best effort traffic can be used as a reference to retrieve metrics from a switch or link, such as packet loss or to estimate the amount of traffic on that route.
 \end{itemize}
-These assumptions indicate that packet loss and path length are part of our cost function: the cost function must find a balance between maintaining low packet loss of the best effort traffic and the delay of QoS flows, caused by using longer paths.\\Furthermore, the Controller must be able to obtain these metrics at any time from any switch in the network before calculating new flow tables.
-This is why we implemented a monitor module on the Controller to gather all network statistics.
-Besides polling for statistics, a switch must also forward events to the Controller when they occur, such as an adjacent link going down or a flow experiencing packet loss.
-These events provide additional monitoring data for the routing algorithms on the Controller.\\Decisions made by the Controller are sent to the switches as updated forwarding tables.
-When the shortest path can still ensure zero packet loss, the Dijkstra shortest path algorithm is executed, which is the default implementation
-  in Floo
- dlight.
+
+These observations show that packet loss and path length are part of our cost function: 
+the cost function must find a balance between maintaining low packet loss of the best effort 
+traffic and the delay of QoS flows, caused by using longer paths.\\
+
+Furthermore, the controller 
+must be able to obtain these metrics at any time from any switch in the network before calculating 
+new flow tables. We implemented a monitor module on the controller to gather all network 
+statistics. Besides polling for statistics, a switch must also forward events to the controller when 
+they occur, such as an link going down or a flow experiencing packet loss. These events provide 
+additional monitoring data for the routing algorithms on the controller.\\
 
 \section{QoS Routing}
-Must ensure zero packet loss base layer, optimise other traffic ... cost function?
+Must ensure zero packet loss base layer, optimize other traffic ... cost function?
 Go from theoretical cost function to exact function in code?
+--> Yes
 
 \section{Evaluation}
-Test scenario (client/sever - filter based on port) + metrics for result + present results\\
+Test scenario (client/server - filter based on port) + metrics for result + present results\\
 An evaluation metric has to represent a value which is relevant to what we want to improve, which is why a measurement must:
@@
 To measure the performance of the network we used the bitrate of all traffic.
-We also show the packet loss of best effort traffic caused by prioritising the QoS flow...
+We also show the packet loss of best effort traffic caused by prioritizing the QoS flow...
 

### 2012-12-06 paper/paper.tex
Paper

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Layer based routing to obtain quality optimization of scalable video streaming in the Future Internet}
+\title{Layer based routing to obtain quality optimization of video streaming in the Future Internet}
 
@@
 This paper presents an OpenFlow controller application which provides 
-Quality of Service (\textit{QoS}) support for Scalable Video Streaming (\textit{SVC}) in an OpenFlow network. 
+Quality of Service (\textit{QoS}) support for video streaming using Scalable Video Coding (\textit{SVC}) in an OpenFlow network. 
 First we route the base layer of SVC encoded videos as a 
@@
 Media applications, such as video streaming, 
-require no packet loss and minimal network delay due to their high Quality of Service (QoS) dependency. 
-When these QoS flows are sent over the same path as other best effort traffic they 
+require no packet loss and minimal network delay. 
+When these video flows are sent over the same path as other best effort traffic they 
 are subject to network congestion and delay which renders an unsatisfactory outcome. 
-Several efforts have been made to support QoS in a TCP/IP-based network, such as DiffServ and IntServ, but these 
+Several efforts have been made to give precedence to certain types of traffic, defined as Quality of Service (QoS).
+There are solutions available for QoS support in a TCP/IP-based network, 
+such as DiffServ \cite{diffserv} and IntServ \cite{intserv}, but these 
 solutions only work well in managed networks.
-Scalable Video Coding (SVC) encodes the stream in a base layer along with several enhancement layers. 
-In order to obtain correct output it is important that the base layer has no 
+Scalable Video Coding (SVC) \cite{schwarz2007overview} encodes the stream in a base layer along with several enhancement layers. 
+In order to obtain qualitative video streams, it is important that the base layer has no 
 packet loss and minimal delay, which is not possible without support for QoS. 
@@
 We also provide failure recovery so the network 
-can converge back to an optimal state when links go down.\\
+can converge back to an optimal state when links fail.\\
 
-We use OpenFlow in order to provide QoS support for scalable video streaming. 
+We use OpenFlow \cite{mckeown2008openflow} in order to provide QoS support for scalable video streaming. 
 OpenFlow is a protocol for switches, allowing to control the packet forwarding process.
@@
 There are many implementations of controllers available for OpenFlow switches.
-The implementation of the controller we will use is Floodlight, because it is open source and widely used.
+The implementation of the controller we will use is Floodlight \cite{floodlight}, because it is open source and widely used.
 
@@
 In our research we identified a need for separate forwarding actions per flow.
-The same principal has been discussed in research papers from Princeton University. 
-In their research the aim was to apply load-balancing during service discovery of web services 
+Wang, R. et al. \cite{wang2011openflow} follow the same principal.
+The aim was to apply load-balancing during service discovery of web services 
 when all requests had to be processed by a central component.
@@
 the use case may be different, the principal is the same as the one applied in our research. 
-When both best effort and QoS flows follow the same paths, chances for congestion and packet loss 
-to occur increase significantly. OpenFlow allows us to set different paths for both flows, which 
+When both best effort and QoS flows follow the same path, chances of congestion and packet loss 
+occurring increase significantly. OpenFlow allows us to set different paths for both flows, which 
 improves the throughput of our QoS flows and also decrease the network load on links used by best 
@@
 The idea to provide a minimal quality for SVC encoded videos originates from 
-research described in [4] and [5]. By routing the base layer as a lossless-QoS flow we can 
+research described in \cite{egilmez2011scalable} and \cite{civanlar2010qos}. 
+By routing the base layer as a lossless-QoS flow we can 
 guarantee a minimal quality to the user. Ideally, the enhancement layers will also be delivered 
 without packet loss or delay but when congestion occurs we demote the enhancement layers to best 
-effort traffic and prioritize the base layer. Research discussed in [4] also attempts to route 
-the enhancement layers as a lossy-QoS flow. Our research goes further and also enables fast failure 
+effort traffic and prioritize the base layer. Research discussed in \cite{egilmez2011scalable} 
+also attempts to route the enhancement layers as a lossy-QoS flow. 
+Our research goes further and also enables fast failure 
 recovery when a link fails.\\
 
-\section{Architecture}
+\section{Design considerations}
 Each OpenFlow network contains at least one controller which is responsible for managing all the switches. 
@@
 
-\begin{figure*}[htb]
+\begin{figure}[htb]
 \centering
@@
 \label{fig_sim}
-\end{figure*}
+\end{figure}
 
 We defined a cost function which can be used to calculate the cost of a path.
-The cost function is based on the following observations. (Insert ref here)
+The cost function is based on the following observations.
 \begin{itemize}
@@
 \bibliographystyle{IEEEtran}
-\bibliography{IEEEabrv,IEEEexample}
+\bibliography{paper}
 

### 2012-12-06 paper/paper.tex
Paper

@@
 high priority QoS flow while the enhancement layers are routed at a lower QoS or as a best-effort flow. 
-Second, we provide failure recovery so the OpenFlow network can converge back to an optimal 
+Secondly, we provide failure recovery so the OpenFlow network can converge back to an optimal 
 state when links fail. We show that using different costs to calculate paths for different flows 
@@
 Media applications, such as video streaming, 
-require no packet loss and minimal network delay. 
+require minimal packet loss and minimal network delay. 
 When these video flows are sent over the same path as other best effort traffic they 
@@
 solutions only work well in managed networks.
+Tunneling with MPLS offers a partial solution, but it lacks real-time reconfigurability and adaptivity. \cite{mpls}\\
+
 Scalable Video Coding (SVC) \cite{schwarz2007overview} encodes the stream in a base layer along with several enhancement layers. 
 In order to obtain qualitative video streams, it is important that the base layer has no 
-packet loss and minimal delay, which is not possible without support for QoS. 
-To solve this issue we need the routing algorithms to detect QoS flows and route these flows differently than 
-normal best effort traffic.
-To achieve this goal in video streaming, we route the base layer of SVC encoded video at a high QoS 
-while the enhancement layers are routed either at a lower QoS or as a best effort flow. 
-We also provide failure recovery so the network 
-can converge back to an optimal state when links fail.\\
+packet loss and minimal delay, which is hard to accomplish without support for QoS. 
+To solve this issue we need to differentiate between the different flows and route these flows 
+differently than normal best effort traffic.\\
 
-We use OpenFlow \cite{mckeown2008openflow} in order to provide QoS support for scalable video streaming. 
+We use OpenFlow \cite{mckeown2008openflow} in order to provide QoS support. 
 OpenFlow is a protocol for switches, allowing to control the packet forwarding process.
@@
 In Section II of this paper we compare related work to our own. 
-In section III we take a closer look at the components of our framework 
-and the principals it is based on. In section IV we present the different algorithms 
-used to determine the behavior of each flow in the network. Section V shows the test 
-scenario we used to evaluate our framework while Section VI presents results.
+In section III we take a closer look at the design considerations for 
+developing an OpenFlow controller application that provides layer-based quality optimization. 
+In section IV we present the general architecture, and in section V we show the algorithms used. 
+Section VI shows experimental results.
 In Section VII we come to a conclusion.
@@
 A problem occurs when each request is forwarded to a centralized component responsible 
-for service discovery, which leads to network congestion. By using OpenFlow, the researchers at 
-Princeton University managed to set separate forwarding actions for client requests which match 
+for service discovery, which leads to network congestion. 
+By using OpenFlow they managed to set separate forwarding actions for client requests which match 
 certain criteria. These actions allow forwarding entities to redirect client requests to an 
@@
 the use case may be different, the principal is the same as the one applied in our research. 
-When both best effort and QoS flows follow the same path, chances of congestion and packet loss 
-occurring increase significantly. OpenFlow allows us to set different paths for both flows, which 
+When all traffic follows the same path, chances of congestion and packet loss 
+occurring increase significantly. OpenFlow allows us to set up different paths for different flows, which 
 improves the throughput of our QoS flows and also decrease the network load on links used by best 
@@
 research described in \cite{egilmez2011scalable} and \cite{civanlar2010qos}. 
-By routing the base layer as a lossless-QoS flow we can 
+By routing the base layer as a lossless QoS flow, which means there is no packet loss, they can 
 guarantee a minimal quality to the user. Ideally, the enhancement layers will also be delivered 
-without packet loss or delay but when congestion occurs we demote the enhancement layers to best 
-effort traffic and prioritize the base layer. Research discussed in \cite{egilmez2011scalable} 
-also attempts to route the enhancement layers as a lossy-QoS flow. 
-Our research goes further and also enables fast failure 
-recovery when a link fails.\\
-
-\section{Design considerations}
-Each OpenFlow network contains at least one controller which is responsible for managing all the switches. 
-A default OpenFlow switch forwards packets 
-to the controller if the packet does not match any forwarding entry. 
-We route SVC encoded video streams as a QoS flow, where we consider the base layer as a 
-high priority QoS flow and the enhancement layer as either a lower QoS flow or as best effort traffic. 
-These principals are shown in Figure \ref{fig_sim}. A filter on the controller separates QoS flows from 
-other traffic, after which the controller can choose a path for the different layers. 
-When the controller calculates a path for a layer, the new forwarding rules are sent to the switches.\\
+without packet loss or delay but when congestion occurs they demote the enhancement layers to best 
+effort traffic and prioritize the base layer.\\
 
@@
 
-We defined a cost function which can be used to calculate the cost of a path.
-The cost function is based on the following observations.
+In our own research we define the base layer as lossless QoS traffic, which means no packet loss is allowed.
+The enhancement layers are lossy QoS traffic, which means they may still follow a different 
+route than best-effort traffic, but some packet loss is allowed.
+These principals are shown in Figure \ref{fig_sim}.
+We also enable fast failure recovery when a link fails, as described in \cite{sharma2011enabling}.
+Also, we present experimental results, whereas other research only presents simulations.\\
+
+\section{Design considerations}
+
+In designing the OpenFlow application, we considered the following.
 \begin{itemize}
@@
 
-These observations show that packet loss and path length are part of our cost function: 
-the cost function must find a balance between maintaining low packet loss of the best effort 
+These observations show that packet loss and path length are important in defining algorithms: 
+the algorithm must find a balance between maintaining low packet loss of the best effort 
 traffic and the delay of QoS flows, caused by using longer paths.\\
 
-Furthermore, the controller 
-must be able to obtain these metrics at any time from any switch in the network before calculating 
+\section{Architecture}
+
+To obtain the different SVC video streams, we use HTTP adaptive streaming.
+This means that the video is split into different segments, and when a new segment starts playing, 
+the following segment is requested using HTTP. 
+If the next segment doesn't arrive in time, it is skipped.
+In the case of SVC streaming the video stream is first split into different layers (base and enhancement layers).
+These layers are then further split into segments of different size, but equal playtime.
+If a segment of a layer does not arrive in time, it is skipped.\\
+
+Each OpenFlow network contains at least one controller which is responsible for managing all the switches. 
+An OpenFlow switch forwards packets 
+to the controller if the packet does not match any forwarding entry in the forwarding table.
+Matching is done on the longest prefix. Wildcards are allowed in the forwarding table.
+A filter in the controller separates QoS flows from 
+other traffic, based on the TCP port, after which the controller can choose a path for the different layers. 
+When the controller calculates a path for a layer, the new forwarding rules are sent to the switches.\\
+
+The different paths are calculated using Dijkstra. 
+Depending on the current load of the network, a cost is given to each link.
+The controller must be able to obtain the load any time from any switch in the network before calculating 
 new flow tables. We implemented a monitor module on the controller to gather all network 
-statistics. Besides polling for statistics, a switch must also forward events to the controller when 
-they occur, such as an link going down or a flow experiencing packet loss. These events provide 
-additional monitoring data for the routing algorithms on the controller.\\
+statistics. Besides polling for statistics, a switch also forwards events to the controller when 
+they occur, such as a link failure. These events are also act upon by the controller, 
+by recalculating a different path for every flow.\\
 
-\section{QoS Routing}
-Must ensure zero packet loss base layer, optimize other traffic ... cost function?
-Go from theoretical cost function to exact function in code?
---> Yes
+\section{Algorithms}
 
-\section{Evaluation}
+For the base layer...
+
+For the enhancement layers...
+
+\section{Experimental results}
+
+All of the experiments are performed on the iLab.t Virtual Wall at iMinds.
+The Virtual Wall is derived from Emulab, developed at Utah university.
+The Virtual Wall facility is a generic test environment for advanced network, 
+distributed software and service evaluation, and supports scalability research.\\
+
+The Virtual Wall facilities consist of 100 nodes 
+(dual processor, dual core servers, 4x1 or 6x1 Gb/s interfaces per node)
+interconnected via a non-blocking 1.5 Tb/s VLAN Ethernet switch enabling flexible creation of experiment topologies, 
+and a display wall (20 monitors) for experiment visualization. 
+Each server is connected with 4 or 6 gigabit Ethernet links to the switch. 
+The Virtual Wall nodes can be assigned different functionalities ranging 
+from terminal, server, network node, and impairment node. 
+Each node is fully dedicated and under full control of one experiment, 
+so the users can test all kind of operating system images and configurations \cite{virtual}.\\
+
 Test scenario (client/server - filter based on port) + metrics for result + present results\\

### 2012-12-10 paper/paper.tex
Update

@@
 %
-%\usepackage{algorithmic}
+\usepackage{algorithmic}
+\usepackage{algorithm}% http://ctan.org/pkg/algorithms
+%\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
 % algorithmic.sty was written by Peter Williams and Rogerio Brito.
@@
 % affiliations
-\author{\IEEEauthorblockN{Piet Smet, Sander Vrijders, Youri Flement, Bruno Chevalier}
+\author{\IEEEauthorblockN{Sander Vrijders, Piet Smet, Youri Flement, Bruno Chevalier}
 \IEEEauthorblockA{Ghent University - iMinds, Department of Information Technology,\\ 
@@
 solutions only work well in managed networks.
-Tunneling with MPLS offers a partial solution, but it lacks real-time reconfigurability and adaptivity. \cite{mpls}\\
+Tunneling with MPLS offers a partial solution, but it lacks real-time reconfigurability and adaptivity. \cite{mpls}
 
@@
 To solve this issue we need to differentiate between the different flows and route these flows 
-differently than normal best effort traffic.\\
+differently than normal best effort traffic.
 
@@
 improves the throughput of our QoS flows and also decrease the network load on links used by best 
-effort traffic.\\
+effort traffic.
 
@@
 without packet loss or delay but when congestion occurs they demote the enhancement layers to best 
-effort traffic and prioritize the base layer.\\
+effort traffic and prioritize the base layer.
 
@@
 \includegraphics[width=\linewidth]{fig/SVC_OpenFlow_flow}
-\caption{Simulation Results}
+\caption{General problem}
 \label{fig_sim}
@@
 We also enable fast failure recovery when a link fails, as described in \cite{sharma2011enabling}.
-Also, we present experimental results, whereas other research only presents simulations.\\
+Also, we present experimental results, whereas other research only presents simulations.
 
@@
 the algorithm must find a balance between maintaining low packet loss of the best effort 
-traffic and the delay of QoS flows, caused by using longer paths.\\
+traffic and the delay of QoS flows, caused by using longer paths.
 
@@
 These layers are then further split into segments of different size, but equal playtime.
-If a segment of a layer does not arrive in time, it is skipped.\\
+If a segment of a layer does not arrive in time, it is skipped.
 
@@
 other traffic, based on the TCP port, after which the controller can choose a path for the different layers. 
-When the controller calculates a path for a layer, the new forwarding rules are sent to the switches.\\
+When the controller calculates a path for a layer, the new forwarding rules are sent to the switches.
 
@@
 they occur, such as a link failure. These events are also act upon by the controller, 
-by recalculating a different path for every flow.\\
+by recalculating a different path for every flow.
 
@@
 
-For the base layer...
+Because the base layer has to be delivered lossless, the weights given to each link depends 
+on the load of the switches connected to these switches.
+The minimum value a link can get is 1, so the hop count is used when the network experiences no load.
+The pseudocode can be seen in Algorithm \ref{alg1}. This algorithm is executed on receiving a packet-in event 
+from a switch with TCP port set to 8080.
 
-For the enhancement layers...
+\begin{algorithm}
+\caption{Calculate route for base layer}
+\label{alg1}
+\begin{algorithmic}
+\FORALL{$link$ in $linksInNetwork$}
+\STATE $leftSwitch \leftarrow link.getSrc()$
+\STATE $rightSwitch \leftarrow link.getDst()$
+\STATE $loadLeft \leftarrow monitorModule.getLoad(LeftSwitch)$
+\STATE $loadRight \leftarrow monitorModule.getLoad(rightSwitch)$
+\STATE $totalLoad \leftarrow loadLeft + loadRight + 1$
+\STATE $linkMap.put(link,weight)$
+\ENDFOR
+$Dijkstra(linkMap)$
+\end{algorithmic}
+\end{algorithm}
+
+For the enhancement layers we followed the same tactic, but we also hold the route that the 
+base layer is following into account. This means we gave a higher cost to the links that the base layer was using, 
+thus ensuring the base layer is delivered lossless. The pseudocode for this algorithm can be seen in Algorithm \ref{alg2}.
+It is executed on packet-in events with the TCP port of the packet set to 8081,8082 or 8083.
+The constant extraLoad can be tweaked to render optimal performance. It must be set high enough to ensure 
+the route the base layer is using is not used, unless the traffic on the other links is much higher.
+
+\begin{algorithm}
+\caption{Calculate route for base layer}
+\label{alg2}
+\begin{algorithmic}
+\FORALL{$link$ in $linksInNetwork$}
+\STATE $leftSwitch \leftarrow link.getSrc()$
+\STATE $rightSwitch \leftarrow link.getDst()$
+\STATE $loadLeft \leftarrow monitorModule.getLoad(LeftSwitch)$
+\STATE $loadRight \leftarrow monitorModule.getLoad(rightSwitch)$
+\STATE $totalLoad \leftarrow loadLeft + loadRight + 1$
+\IF{$linksRouteBaseLayer.contains(link)$}
+\STATE $totalLoad \leftarrow totalLoad + Constants.extraLoad()$
+\ENDIF
+\STATE $linkMap.put(link,weight)$
+\ENDFOR
+$Dijkstra(linkMap)$
+\end{algorithmic}
+\end{algorithm}
 
@@
 The Virtual Wall facility is a generic test environment for advanced network, 
-distributed software and service evaluation, and supports scalability research.\\
+distributed software and service evaluation, and supports scalability research.
 
@@
 Each node is fully dedicated and under full control of one experiment, 
-so the users can test all kind of operating system images and configurations \cite{virtual}.\\
+so the users can test all kind of operating system images and configurations \cite{virtual}.
 
-Test scenario (client/server - filter based on port) + metrics for result + present results\\
+We created the topology shown in Figure \ref{fig:topology}, on which we performed the experiments.
+
+\begin{figure}[htb]
+\centering
+\includegraphics[scale=0.4]{fig/topology}
+\caption{Topology used}
+\label{fig:topology}
+\end{figure}
+
 An evaluation metric has to represent a value which is relevant to what we want to improve, which is why a measurement must:

### 2012-12-11 floodlight/src/main/java/net/floodlightcontroller/core/NetworkStatistics.java
statistieken met capaciteit

@@ -101,9 +101,10 @@ public class NetworkStatistics extends TimerTask {
 		    		newload = reply.getTransmitBytes();		
 		    		loadHistory.put(sw.getId(),newload);
 		    	}				    	
-		    	
-		    	currentLoad.put(sw.getId(), newload / (period / 1000));
-		    	//System.out.println("Load for switch: " + currentLoad.get(sw) + " bytes/second");
+		    	//System.out.println("Capa: " + sw.getCapabilities() * 1024);
+		    	//System.out.println("Load: " + newload);
+		    	currentLoad.put(sw.getId(), (sw.getCapabilities() * 1024) - newload);
+		    	//System.out.println("Load for switch: " + currentLoad.get(sw.getId()));
 		    }
 		}
 	}
@@ -118,7 +119,8 @@ public class NetworkStatistics extends TimerTask {
 			//System.out.println("Loads in map:" + currentLoad.size());
 			return currentLoad;
 		} else {
-			return null;
+			//niet goed, moet null terug geven, eerst aanpassen in topo instance
+			return currentLoad;
 		}		
 	}
 	

### 2012-12-11 SVC-Client/src/svc/client/Constants.java
Constants

@@ -11,8 +11,8 @@ package svc.client;
 public class Constants {
     //Set constants
 
-    public static final int NR_SEGMENTS = 3;
-    public static final long MILLISECONDS = 4024;
-    public static final String SERVER_LOCATION = "google.com";
+    public static final int NR_SEGMENTS = 77;
+    public static final long MILLISECONDS = 4266;
+    public static final String SERVER_LOCATION = "10.1.3.2";
     public static final int[] PORTS = {8080, 8081, 8082, 8083};
 }

### 2012-12-11 floodlight/src/main/java/net/floodlightcontroller/core/NetworkStatistics.java
added write to file

@@ -8,6 +8,7 @@ import java.util.List;
 import java.util.Map;
 import java.util.TimerTask;
 import java.util.concurrent.Future;
+import java.io.*;
 
 import org.openflow.protocol.OFPort;
 import org.openflow.protocol.OFStatisticsRequest;
@@ -35,6 +36,10 @@ public class NetworkStatistics extends TimerTask {
 	
 	//a map that keeps record of the current load
 	protected static Map<Long, Long> currentLoad;
+
+	private FileWriter fw;
+	private BufferedWriter bw;
+	private static int statsCounter = 0;
 	
 	/**
 	 * Creates a new statistics task
@@ -46,6 +51,7 @@ public class NetworkStatistics extends TimerTask {
 		this.period = period;
 		loadHistory = new HashMap<Long, Long>();
 		currentLoad = new HashMap<Long, Long>();
+		statsCounter++;
 	}
 	
 	/**
@@ -57,55 +63,64 @@ public class NetworkStatistics extends TimerTask {
 		Map<Long,IOFSwitch> switches = floodlightProvider.getSwitches();				
 		Collection<IOFSwitch> swList = switches.values();		
 		Iterator<IOFSwitch> it = swList.iterator();
-		
-		//for each switch, retrieve the statistics
-		while(it.hasNext()){
-			IOFSwitch sw = it.next();
-			List<OFStatistics> stats = new ArrayList<OFStatistics>();
+		try{
+			fw = new FileWriter("netstatslog_"+statsCounter+".txt",true);
+			bw = new BufferedWriter(fw);
 			
-			//make a statistics request			
-		    OFStatisticsRequest req = new OFStatisticsRequest();
-		    req.setStatisticType(OFStatisticsType.PORT);
-            req.setXid(sw.getNextTransactionId());
-		    
-            //fill it with a specific request
-		    OFPortStatisticsRequest specReq = new OFPortStatisticsRequest();
-		    specReq.setPortNumber(OFPort.OFPP_NONE.getValue());
-		    stats.add((OFStatistics)specReq);
-		    req.setStatistics(stats);
-		    req.setLengthU(req.getLengthU() + specReq.getLength());
-		    
-		    //attempt to retrieve the statistics
-		    Future<List<OFStatistics>> future = null;
-		    List<OFStatistics> values = null;
-		    
-		    try {
-		    	future = sw.getStatistics(req);
-		    	values = future.get();		    	
-		    } catch (Exception e){
[…]

### 2012-12-11 floodlight/src/main/java/net/floodlightcontroller/core/NetworkStatistics.java
fixed controller

@@ -64,7 +64,7 @@ public class NetworkStatistics extends TimerTask {
 		Collection<IOFSwitch> swList = switches.values();		
 		Iterator<IOFSwitch> it = swList.iterator();
 		try{
-			fw = new FileWriter("netstatslog_"+statsCounter+".txt",true);
+			fw = new FileWriter("netstatslog.txt",true);
 			bw = new BufferedWriter(fw);
 			
 			//for each switch, retrieve the statistics
@@ -103,7 +103,7 @@ public class NetworkStatistics extends TimerTask {
 			    	long newload = 0;		    	
 			    	if(currentLoad.containsKey(sw.getId())){
 			    		//we've seen the switch before
-			    		newload = reply.getTransmitBytes() - loadHistory.get(sw.getId());
+			    		newload = reply.getTransmitBytes();
 				    	loadHistory.put(sw.getId(), reply.getTransmitBytes());  		
 			    	} else {
 			    		//this is the first time we see the switch
@@ -113,11 +113,13 @@ public class NetworkStatistics extends TimerTask {
 			    	//System.out.println("Capa: " + sw.getCapabilities() * 1024);
 			    	//System.out.println("Load: " + newload);
 			    	currentLoad.put(sw.getId(), (sw.getCapabilities() * 1024) - newload);
-			    	bw.write("id: "+sw.getId()+" ;load: "+((sw.getCapabilities() * 1024) - newload));
+			    	bw.write("\n id: "+sw.getId()+" ;load: "+((sw.getCapabilities() * 1024) - newload));
+			    	
 			    	//System.out.println("Load for switch: " + currentLoad.get(sw.getId()));
 			    }
 			}
-
+			bw.write("\n******************");
+			
 			bw.close();
 		}catch (Exception e){//Catch exception if any
 			System.err.println("Error: " + e.getMessage());
@@ -135,7 +137,7 @@ public class NetworkStatistics extends TimerTask {
 			return currentLoad;
 		} else {
 			//niet goed, moet null terug geven, eerst aanpassen in topo instance
-			return currentLoad;
+			return null;
 		}		
 	}
 	

### 2012-12-11 floodlight/src/main/java/net/floodlightcontroller/topology/TopologyInstance.java

@@ -647,75 +647,10 @@ public class TopologyInstance {
 	//	if (tcpport != 0) {
 			// TODO: Construct a map with link weights
 			//indexen via switch ID
+		if (tcpport != 5001 && tcpport != 100) {
+			/*
 			Map<Long, Long> load = NetworkStatistics.networkLoad();
 			Map<Link, Integer> linkCost = new HashMap<Link, Integer>();
-
-			if (tcpport == 8080) {
-				for (NodePortTuple nt : tunnelPorts) {
-					if (switchPortLinks.get(nt) == null)
-						continue;
-					for (Link link : switchPortLinks.get(nt)) {
-						if (link == null)
-							continue;
-
-						// Here you have the link, now add a weight
-						long src = link.getSrc();
-						long dst = link.getDst();
-						
-						long srcLoad = load.get(src);
-						long dstLoad = load.get(dst);
-
-						// SAFE?
-						int weight = (int) (srcLoad + dstLoad) + 1;
-
-						linkCost.put(link, weight);
-					}
-				}
-			} else if (tcpport == 8081) {
-				for (NodePortTuple nt : tunnelPorts) {
-					if (switchPortLinks.get(nt) == null)
-						continue;
-					for (Link link : switchPortLinks.get(nt)) {
-						if (link == null)
-							continue;
-
-						// Here you have the link, now add a weight
-						long src = link.getSrc();
-						long dst = link.getDst();
-
-						long srcLoad = load.get(src);
-						long dstLoad = load.get(dst);
-
-						// SAFE?
-						int weight = (int) (srcLoad + dstLoad) + 1;
-
-						linkCost.put(link, weight);
-					}
-				}
-			} else if (tcpport == 8082) {
-				for (NodePortTuple nt : tunnelPorts) {
-					if (switchPortLinks.get(nt) == null)
-						continue;
-					for (Link link : switchPortLinks.get(nt)) {
-						if (link == null)
-							continue;
-
-						// Here you have the link, now add a weight
-						long src = link.getSrc();
-						long dst = link.getDst();
-
-						long srcLoad = load.get(src);
-						long dstLoad = load.get(dst);
-
-						// SAFE?
-						int weight = (int) (srcLoad + dstLoad) +1;
-
-						linkCost.put(link, weight);
-					}
-				}
-			//} else if (tcpport == 8083) {
-			} else { 
-				
 				
[…]

### 2012-12-11 scripts/topo.py
Update

@@ -1,0 +1,48 @@
+"""Topology for the ODS project
+
+author: Sander Vrijders
+
+Adding the 'ofsvc' dict with a key/value pair to generate our newly defined topology enables one to pass in '--topo=ofsvc' from the command line. """
+
+from mininet.topo import Topo, Node
+
+class MyTopo( Topo ):
+
+    def __init__( self, enable_all = True ):
+        "Create custom topo."
+
+        # Add default members to class.
+        super( MyTopo, self ).__init__()
+
+        # Set Node IDs for hosts and switches
+        src = 1
+        switch0 = 2
+        switch1 = 3
+        switch2 = 4
+        switch3 = 5
+        switch4 = 6
+        sink = 7
+
+        # Add nodes
+        self.add_node( src, Node( is_switch=False ) )
+        self.add_node( switch0, Node( is_switch=True ) )
+        self.add_node( switch1, Node( is_switch=True ) )
+        self.add_node( switch2, Node( is_switch=True ) )
+        self.add_node( switch3, Node( is_switch=True ) )
+        self.add_node( switch4, Node( is_switch=True ) )
+        self.add_node( sink, Node( is_switch=False ) )
+
+        # Add edges
+        self.add_edge( src, switch0 )
+        self.add_edge( switch0, switch1 )
+        self.add_edge( switch0, switch2 )
+        self.add_edge( switch1, switch2 )
+        self.add_edge( switch1, switch3 )
+        self.add_edge( switch2, switch4 )
+        self.add_edge( switch3, switch4 )
+        self.add_edge( switch4, sink )  
+
+        # Consider all switches and hosts 'on'
+        self.enable_all()
+
+topos = { 'ofsvc': ( lambda: MyTopo() ) } 

### 2012-12-11 openv/installopenv.sh
Update

@@ -1,0 +1,12 @@
+#!/bin/bash
+sudo apt-get update
+sudo apt-get install gawk
+
+cd ~/openv/openvswitch-1.1.0pre2
+sudo /sbin/insmod ./datapath/linux-2.6/openvswitch_mod.ko
+./getEthernetPortsSpace.sh | sudo xargs ./utilities/ovs-dpctl add-dp dp0
+sudo ifconfig dp0 up
+sudo mkdir /usr/local/var
+sudo mkdir /usr/local/var/run
+sudo mkdir /usr/local/var/run/openvswitch
+sudo ./utilities/ovs-openflowd dp0 tcp:10.1.8.2 --pidfile=$$

### 2012-12-11 openv/openvswitch-1.1.0pre2/config.h

@@ -1,0 +1,155 @@
+/* config.h.  Generated from config.h.in by configure.  */
+/* config.h.in.  Generated from configure.ac by autoheader.  */
+
+/* Define if building universal (internal helper macro) */
+/* #undef AC_APPLE_UNIVERSAL_BUILD */
+
+/* Official build number as a VERSION suffix string, e.g. "+build123", or ""
+   if this is not an official build. */
+#define BUILDNR ""
+
+/* Define to 1 if net/if_packet.h is available. */
+#define HAVE_IF_PACKET 1
+
+/* Define to 1 if you have the <inttypes.h> header file. */
+#define HAVE_INTTYPES_H 1
+
+/* Define to 1 if you have the `socket' library (-lsocket). */
+/* #undef HAVE_LIBSOCKET */
+
+/* Define to 1 if you have __malloc_hook, __realloc_hook, and __free_hook in
+   <malloc.h>. */
+#define HAVE_MALLOC_HOOKS 1
+
+/* Define to 1 if you have the <memory.h> header file. */
+#define HAVE_MEMORY_H 1
+
+/* Define to 1 if you have the `mlockall' function. */
+#define HAVE_MLOCKALL 1
+
+/* Define to 1 if Netlink protocol is available. */
+#define HAVE_NETLINK 1
+
+/* Define to 1 if OpenSSL is installed. */
+#define HAVE_OPENSSL 1
+
+/* Define to 1 if you have the <stdint.h> header file. */
+#define HAVE_STDINT_H 1
+
+/* Define to 1 if you have the <stdlib.h> header file. */
+#define HAVE_STDLIB_H 1
+
+/* Define to 1 if you have the <strings.h> header file. */
+#define HAVE_STRINGS_H 1
+
+/* Define to 1 if you have the <string.h> header file. */
+#define HAVE_STRING_H 1
+
+/* Define to 1 if you have the `strsignal' function. */
+#define HAVE_STRSIGNAL 1
+
+/* Define if strtok_r macro segfaults on some inputs */
+/* #undef HAVE_STRTOK_R_BUG */
+
+/* Define to 1 if `st_mtimensec' is a member of `struct stat'. */
+/* #undef HAVE_STRUCT_STAT_ST_MTIMENSEC */
+
+/* Define to 1 if `st_mtim.tv_nsec' is a member of `struct stat'. */
+#define HAVE_STRUCT_STAT_ST_MTIM_TV_NSEC 1
+
[…]

### 2012-12-11 openv/openvswitch-1.1.0pre2/datapath/linux-2.6/brcompat_mod.mod.c

@@ -1,0 +1,71 @@
+#include <linux/module.h>
+#include <linux/vermagic.h>
+#include <linux/compiler.h>
+
+MODULE_INFO(vermagic, VERMAGIC_STRING);
+
+struct module __this_module
+__attribute__((section(".gnu.linkonce.this_module"))) = {
+ .name = KBUILD_MODNAME,
+ .init = init_module,
+#ifdef CONFIG_MODULE_UNLOAD
+ .exit = cleanup_module,
+#endif
+ .arch = MODULE_ARCH_INIT,
+};
+
+static const struct modversion_info ____versions[]
+__used
+__attribute__((section("__versions"))) = {
+	{ 0xae141548, "module_layout" },
+	{ 0xdca0e950, "genl_register_family" },
+	{ 0xab910849, "single_open" },
+	{ 0xd0d8621b, "strlen" },
+	{ 0x49439411, "genl_unregister_family" },
+	{ 0xe4b7d1f2, "single_release" },
+	{ 0xe66df013, "seq_puts" },
+	{ 0xc7a4fbed, "rtnl_lock" },
+	{ 0x901e58ed, "skb_clone" },
+	{ 0x2bc801eb, "remove_proc_entry" },
+	{ 0x446da885, "mutex_unlock" },
+	{ 0x712aa29b, "_spin_lock_irqsave" },
+	{ 0x19fae545, "dp_ioctl_hook" },
+	{ 0x1f33bd3a, "seq_read" },
+	{ 0xc499ae1e, "kstrdup" },
+	{ 0xe2d5255a, "strcmp" },
+	{ 0x5ce21240, "proc_mkdir" },
+	{ 0xb86e4ab9, "random32" },
+	{ 0xb72397d5, "printk" },
+	{ 0x5152e605, "memcmp" },
+	{ 0x2da418b5, "copy_to_user" },
+	{ 0x4307b811, "nla_put" },
+	{ 0x16305289, "warn_slowpath_null" },
+	{ 0xc7423ae, "mutex_lock" },
+	{ 0x4b07e779, "_spin_unlock_irqrestore" },
+	{ 0xa0068f7b, "netlink_unicast" },
+	{ 0xf6b374cb, "init_net" },
+	{ 0xca49a0b0, "__dev_get_by_index" },
+	{ 0x95cbadbf, "__alloc_skb" },
+	{ 0xfbf219b, "netlink_broadcast" },
[…]

### 2012-12-11 openv/openvswitch-1.1.0pre2/datapath/linux-2.6/openvswitch_mod.mod.c

@@ -1,0 +1,193 @@
+#include <linux/module.h>
+#include <linux/vermagic.h>
+#include <linux/compiler.h>
+
+MODULE_INFO(vermagic, VERMAGIC_STRING);
+
+struct module __this_module
+__attribute__((section(".gnu.linkonce.this_module"))) = {
+ .name = KBUILD_MODNAME,
+ .init = init_module,
+#ifdef CONFIG_MODULE_UNLOAD
+ .exit = cleanup_module,
+#endif
+ .arch = MODULE_ARCH_INIT,
+};
+
+static const struct modversion_info ____versions[]
+__used
+__attribute__((section("__versions"))) = {
+	{ 0xae141548, "module_layout" },
+	{ 0x87ba0fb7, "register_netdevice" },
+	{ 0x1a0acb7e, "kobject_put" },
+	{ 0x6bc3fbc0, "__unregister_chrdev" },
+	{ 0xd6c23881, "kmem_cache_destroy" },
+	{ 0xa4709493, "per_cpu__current_task" },
+	{ 0xfc2358d1, "kmalloc_caches" },
+	{ 0x12da5bb2, "__kmalloc" },
+	{ 0x9b388444, "get_zeroed_page" },
+	{ 0x511b81fe, "skb_gso_segment" },
+	{ 0xda3863ca, "inet_frag_kill" },
+	{ 0x1c1af916, "set_normalized_timespec" },
+	{ 0x57b57ebe, "jiffies_to_timespec" },
+	{ 0xf100841c, "unregister_netdevice" },
+	{ 0xd0d8621b, "strlen" },
+	{ 0x1a75caa3, "_read_lock" },
+	{ 0xf1c8e6fc, "ip_route_output_key" },
+	{ 0xc0831457, "dev_get_flags" },
+	{ 0x60a13e90, "rcu_barrier" },
+	{ 0xf568fbe3, "dev_disable_lro" },
+	{ 0xc8b57c27, "autoremove_wake_function" },
+	{ 0x5bf7330e, "ip_local_out" },
+	{ 0x79aa04a2, "get_random_bytes" },
+	{ 0xc7a4fbed, "rtnl_lock" },
+	{ 0xca8c759, "sock_release" },
[…]

### 2012-12-11 openv/openvswitch-1.1.0pre2/getEthernetPortsSpace.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+ifconfig | gawk -f ~/openv/openvswitch-1.1.0pre2/getEthernetPortsSpace.gawk

### 2012-12-11 openv/openvswitch-1.1.0pre2/lib/coverage-counters.c

@@ -1,0 +1,392 @@
+#include "coverage-counters.h"
+#include <stddef.h>
+#include "coverage.h"
+#include "util.h"
+
+/* vswitchd/bridge.c:1204 */
+struct coverage_counter bridge_flush_count = { "bridge_flush", 0, 0 };
+
+/* vswitchd/bridge.c:2521 */
+struct coverage_counter bridge_process_flow_count = { "bridge_process_flow", 0, 0 };
+
+/* vswitchd/bridge.c:555 */
+struct coverage_counter bridge_reconfigure_count = { "bridge_reconfigure", 0, 0 };
+
+/* lib/dpif.c:384 */
+struct coverage_counter dpif_destroy_count = { "dpif_destroy", 0, 0 };
+
+/* lib/dpif.c:934 */
+struct coverage_counter dpif_execute_count = { "dpif_execute", 0, 0 };
+
+/* lib/dpif.c:825 */
+struct coverage_counter dpif_flow_del_count = { "dpif_flow_del", 0, 0 };
+
+/* lib/dpif.c:697 */
+struct coverage_counter dpif_flow_flush_count = { "dpif_flow_flush", 0, 0 };
+
+/* lib/dpif.c:722, lib/dpif.c:774 */
+struct coverage_counter dpif_flow_get_count = { "dpif_flow_get", 0, 0 };
+
+/* lib/dpif.c:806 */
+struct coverage_counter dpif_flow_put_count = { "dpif_flow_put", 0, 0 };
+
+/* lib/dpif.c:850 */
+struct coverage_counter dpif_flow_query_list_count = { "dpif_flow_query_list", 0, 0 };
+
+/* lib/dpif.c:866 */
+struct coverage_counter dpif_flow_query_list_n_count = { "dpif_flow_query_list_n", 0, 0 };
+
+/* lib/dpif.c:443 */
+struct coverage_counter dpif_port_add_count = { "dpif_port_add", 0, 0 };
+
+/* lib/dpif.c:467 */
+struct coverage_counter dpif_port_del_count = { "dpif_port_del", 0, 0 };
+
+/* lib/dpif.c:683 */
+struct coverage_counter dpif_port_group_set_count = { "dpif_port_group_set", 0, 0 };
+
+/* lib/dpif.c:1064 */
+struct coverage_counter dpif_purge_count = { "dpif_purge", 0, 0 };
+
+/* lib/flow.c:148 */
+struct coverage_counter flow_extract_count = { "flow_extract", 0, 0 };
+
[…]

### 2012-12-11 openv/openvswitch-1.1.0pre2/lib/dirs.c

@@ -1,0 +1,5 @@
+/* -*- mode: c; buffer-read-only: t -*- */
+const char ovs_pkgdatadir[] = "/usr/local/share/openvswitch";
+const char ovs_rundir[] = "/usr/local/var/run/openvswitch";
+const char ovs_logdir[] = "/usr/local/var/log/openvswitch";
+const char ovs_bindir[] = "/usr/local/bin";

### 2012-12-11 scripts/getEthernetPorts.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+
+ifconfig | gawk -f ./getEthernetPorts.gawk
+

### 2012-12-11 scripts/installopenvswitch.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+
+#see http://networkstatic.net/openvswitch-configure-from-packages-and-attaching-to-a-floodlight-openflow-controller/
+sudo apt-get update
+sudo apt-get purge network-manager
+sudo apt-get install debhelper
+sudo apt-get install module-assistant
+sudo apt-get install uuid-runtime
+sudo apt-get install bridge-utils
+sudo apt-get install python-argparse
+sudo dpkg -i openvswitch-common_1.9.90-1_i386.deb
+sudo dpkg -i openvswitch-switch_1.9.90-1_i386.deb
+sudo dpkg -i openvswitch-brcompat_1.9.90-1_i386.deb
+sudo dpkg -i openvswitch-pki_1.9.90-1_all.deb
+sudo dpkg -i openvswitch-controller_1.9.90-1_i386.deb
+sudo dpkg -i openvswitch-switch_1.9.90-1_i386.deb
+sudo dpkg -i openvswitch-datapath-source_1.9.90-1_all.deb
+
+sudo module-assistant auto-install openvswitch-datapath
+
+

### 2012-12-11 scripts/ofcontroller.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+(( $# == 0 )) || { echo "usage: ./ofcontroller.sh"; exit 1; }
+sudo ofprotocol unix:/var/run/dp0 tcp:10.1.8.2:6633

### 2012-12-11 scripts/ofswitch.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+(( $# == 1 )) || { echo "usage: ./ofswitch [path to csv of interfaces]"; exit 1; }
+sudo ofdatapath --detach punix:/var/run/dp0 -i $(cat $1)
+

### 2012-12-11 scripts/startclient.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+java -jar SVC-Client/dist/SVC-Client.jar

### 2012-12-11 scripts/startfloodlight.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+cd ~
+java -jar ./floodlight/target/floodlight.jar

### 2012-12-11 scripts/stop_java.sh

@@ -1,0 +1,10 @@
+#!/bin/sh
+
+node_count=32
+
+for node_id in `jot $node_count 1`; do
+	hostname="peer-$node_id"
+    ssh $hostname.group16.odspracticum.wall.test "killall java" &
+done
+
+wait

### 2012-12-11 scripts/switchofdatapath.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+./getEthernetPorts.sh | xargs sudo ofdatapath --detach punix:/var/run/dp0 -i
+
+

### 2012-12-11 scripts/switchscript.sh

@@ -1,0 +1,12 @@
+#!/bin/bash                                                                                                                                                      
+cd ~/openflow-1.0.0                                                                                                    
+sudo make install                                                                                                                                                
+sudo apt-get update && sudo apt-get install emacs xauth --yes
+
+sudo apt-get install gawk
+sudo update-alternatives --set awk /usr/bin/gawk
+alias ll='ls -al'
+
+cd ~
+
+./getEthernetPorts.sh | xargs sudo ofdatapath --detach punix:/var/run/dp0 -i

### 2012-12-11 SVC-Client/src/svc/client/GetSegment.java
uopdate

@@ -22,12 +22,14 @@ public class GetSegment extends Thread {
     private final HttpContext context;
     private final HttpGet httpget;
     private final int id;
+    private final int segment;
 
-    public GetSegment(HttpClient httpClient, HttpGet httpget, int id) {
+    public GetSegment(HttpClient httpClient, HttpGet httpget, int id, int segment) {
         this.httpClient = httpClient;
         this.context = new BasicHttpContext();
         this.httpget = httpget;
         this.id = id;
+        this.segment = segment;
     }
 
     public HttpGet getHttpget() {
@@ -38,7 +40,7 @@ public class GetSegment extends Thread {
 
     @Override
     public synchronized void run() {
-        System.out.println(id + " - about to get something from " + httpget.getURI());
+        //System.out.println(id + " - about to get something from " + httpget.getURI());
 
         try {
          
@@ -50,7 +52,8 @@ public class GetSegment extends Thread {
             HttpEntity entity = response.getEntity();
             if (entity != null) {
                 byte[] bytes = EntityUtils.toByteArray(entity);
-                System.out.println(id + " - " + httpget.getURI() + " - " + bytes.length + " bytes read");
+                //System.out.println(id + " - " + httpget.getURI() + " - " + bytes.length + " bytes read");
+                System.out.println(segment + " " + id);
             }
 
         } catch (Exception e) {

### 2012-12-11 SVC-Client/src/svc/client/SVCClient.java

@@ -34,7 +34,7 @@ public class SVCClient {
                     //Location ~= serverlocatie:8080/video_s0_l0.dat
                     String location = "http://" + Constants.SERVER_LOCATION + ":" + Constants.PORTS[i] + "/video_s" + k + "_l" + i + ".dat";
                     HttpGet httpget = new HttpGet(location);
-                    threads[i] = new GetSegment(httpclient, httpget, i);
+                    threads[i] = new GetSegment(httpclient, httpget, i,k);
                 }
 
 

### 2012-12-11 paper/paper.tex
mja

@@
 We created the topology shown in Figure \ref{fig:topology}, on which we performed the experiments.
+Each link has a 1 Gb connection. The switches are running OpenVSwitch instead of the reference user-space switch, 
+because Floodlight won't work with the reference user-space switch.
 
@@
 
-An evaluation metric has to represent a value which is relevant to what we want to improve, which is why a measurement must:
-\begin{itemize}
-	\item represent the quality of the received video
-	\item represent the performance of the network
-	\item represent the impact on best effort traffic
-\end{itemize}
-To measure quality of the video we have chosen for the PSNR value of the received video compared to the raw footage.
+To measure the quality of the video we chose the PSNR value of the received video compared to the raw footage. 
 The higher the PSNR value, the better the frames were preserved during transfer. 
-To measure the performance of the network we used the bitrate of all traffic.
-We also show the packet loss of best effort traffic caused by prioritizing the QoS flow...
+A base layer has a low PSNR value, and every extra enhancement layer that is also received, 
+increases this value. %In Figure \ref{fig:psnr} we show the PSNR value as a function of time.
+Unfortunately, we did not have enough time to obtain results.
 
 \section{Conclusion}
-The conclusion goes here.
+In this paper we described a way to perform quality optimization of video streaming.
+We do this by using HTTP adaptive streaming of an SVC encoded video stream, which is composed of a base layer and three enhancement layers.
+Each layer is routed differently, depending on the required Quality of Service. 
+The base layer is delivered as a lossless video stream. This means it is always delivered. It has the highest QoS priority.
+The enhancement layers are considered lossy video streams, which means they may not always arrive, 
+depending on the network load. We also implemented fast failure recovery. This means if a link fails, the routes get recalculated very fast, 
+rendering an optimal performance.
+
+In future work we will try to optimize the weights given to each link in the algorithm. 
+This will give an even higher quality optimization.
 
